---
title: "Analizing Protests from 1990 to 2020"
author: "Angel Feliz"
date: "`r format(Sys.Date(),'%d %B %Y')`"
output: 
  html_document:
    css: styles.css
    code_folding: show
    highlight: tango
    theme: united
    toc: true
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.align = "center", fig.dim = c(12,8))
```

# Introduction

Protests are an important part of society change. In this post we are going to explore data from **The Mass Mobilization Project** which contains information about protests against governments, and contains a variety of variables such as location, dates and number of participants. The data was obtained by this [link](https://github.com/datacamp/careerhub-data/tree/master/Mass%20Protest%20Data)

This data opens the possibility to understand:

1. Where do protest occurs?
2. Why do people protest?
3. How do states react to protest?
4. Does a good state response increase the number protests?
5. What does characterize an protest? (number of demands)
6. Which kind of people participate in protests?

To answer those questions we will perform an **Exploratory Data Analysis (EDA)** applying the next skills:

- R (functional) programming with **base R**
- Data manipulation with **data.table** (much faster than dplyr of tidyverse)
- Date manipulation with **lubridate** part of tidyverse
- String manipulation with **stringr** part of tidyverse
- Data visualization with **ggplot2**, **scales** and **forcats** parts of tidyverse
- Interactive data visualization with **plotly** and **crosstalk**
- Automated reporting with **rmarkdown**, **flextable**, **html** and **css**
- Text mining with **tidytext**


```{r initial-set-up, class.source = 'fold-hide'}

# 1. Libraries importation
library(data.table)
library(lubridate)
library(stringr)
library(countrycode)
library(flextable)
library(ggplot2)
library(scales)
library(forcats)
library(plotly)
library(crosstalk)
theme_set(theme_light())


# 2. Custom vectors

EuropeMissingCountries <-  
c("Yugoslavia", "Serbia and Montenegro",
  "Kosovo","Germany East","Czechoslovakia")


# Next variables help to extract the number of 
# participants from participants columns.

KeyWordsAfterNumber <- 
str_c(c("people","protesters","drivers","residents","supporters",
        "members","participants","former","demonstrators"),
      collapse = "|")

KeyWordsBeforeNumber <-
str_c(c("about","around","more than",">","<","almost","over","more than"),
      collapse = "|")


# 3. Custom function

# plot_chr_count
# It counts the unique number of id_var for categories included in group_var
# and plot_fill_var a col plot.
#   
#  Arguments
#  DATA: A data.table with data to count.
#  id_var: A character value that defines the id column to count the events.
#  group_var: A character value that defines the column to group by.
#  plot_fill_var: A character value that defines the column to group by
#                 and fill the plot.
#  plot_fill: A character value with constant color to fill all bars.
#  plot_title: A character value that defines the plot title.
#  plot_max: A numeric value that defines the number of categories 
#            that will appear in the plot.
#  use_cumulative_sum: A boolean value to perform a cumulative sum of percentages.

plot_chr_count <- function(DATA,
                           id_var,
                           group_var,
                           plot_fill_var,
                           plot_fill,
                           plot_title,
                           plot_max,
                           use_cumulative_sum = TRUE){
  
  
  # To be able to color bar according to an extra variable it's needed
  # that all the data to be summarize by the color variable.
  
  if(!missing(plot_fill_var)){
    group_vector <- c("group",plot_fill_var)
  }else{
    group_vector <- "group"
  }
  
  
  data_summary <-
    DATA[, setnames(.SD, c(id_var,group_var), c("id","group"))
    ][, .(number_protest = uniqueN(id)), group_vector
    ][order(-number_protest)
    ][, group := as_factor(group)
    ][, pct := fifelse(rep(use_cumulative_sum, .N), 
                       cumsum(number_protest), number_protest) /
               sum(number_protest)]
  
  
  # Filtering top categories
  
  if(!missing(plot_max) && !nrow(data_summary) <= plot_max){
    data_summary <- data_summary[1:plot_max]
  }
  
  
  # Defining data plot and plot title 
  
  if(missing(plot_fill_var) & missing(plot_fill)){
    
    base_aes <- 
      ggplot(data_summary,aes(number_protest, fct_rev(group))) +
      geom_col(aes(fill = group))
    
    fill_title <- str_to_sentence(group_var)
    
  }else if(!missing(plot_fill)){
    
    base_aes <-
      ggplot(data_summary,aes(number_protest, fct_rev(group))) +
      geom_col(fill = plot_fill)
    
    fill_title <- ""
    
  }else{
    
    data_summary[, plot_fill_var := get(plot_fill_var)]
    base_aes <-
      ggplot(data_summary,aes(number_protest, fct_rev(group))) +
      geom_col(aes(fill = plot_fill_var))
    
    fill_title <-str_to_sentence(plot_fill_var)
    
  }
  
  
  # Applying theme customization
  
  base_aes +
    geom_text(aes(label = percent(pct, accuracy = 0.1)), hjust = -0.3, size = 6)+
    geom_blank(aes(x = number_protest *1.08))+
    scale_x_continuous(labels = comma_format(accuracy = 1))+
    labs(title = plot_title, x = "", y = "", fill = fill_title)+
    theme(plot.title = element_text(color = "black", face = "bold", 
                                    size = 22, hjust = 0.5),
          legend.position = "top",
          legend.title = element_text(color = "black",size = 16),
          legend.text = element_text(color = "black",size = 14),
          axis.title = element_text(color = "black", face = "bold", size = 18),
          axis.text = element_text(color = "black", size = 16),
          strip.text = element_text(face = "bold",size = 14),
          panel.grid.major.y = element_blank())
}

```


# 1. Data wrangling

Before being able to explore the data it's important to face some problems with the data. The next table summarizes the problems founds in the data and the problems to solve each problem:

<table class = "wrangling-table">
 <thead>
  <tr>
    <th>Order</th> 
    <th>Problems</th> 
    <th>Solutions</th> 
  </tr>
 </thead>
 <tbody>
  <tr>
    <td>01</td> 
    <td>Many rows don't represent a protest to analyze.</td> 
    <td>We took those rows off.</td> 
  </tr>
  <tr>
    <td>02</td> 
    <td>Days, months and years were stored in individual columns to define protest start and end dates.</td> 
    <td>We created Date variables to define the protest start and end dates.</td> 
  </tr>
  <tr>
    <td>03</td> 
    <td>`participants_category` column has a small number of categories but it has many missing values and `participants` mixes value with words</td> 
    <td>We cleaned `participants` and then complete the `participants_category` with that information.</td> 
   </tr>
    <tr>
    <td>04</td> 
    <td>Protester demands and state responses were stored in many columns as one protest can have many values in both situation</td> 
    <td>We reshape the data to have longer table where protester demands and state responses are saved in 3 columns without missing values.</td> 
   </tr>
   
  </tbody>
</table>

Below we can see a protest `id` as an example of the changes made. That table has the next characteristics:

1. Most of the protests have a `participants_range` category.
2. All protests have a`start_date` and an `end_date` in a Date format.
3. Every protest has a row for every possible action that could occur during the protest ether made by protesters or the state and the last column confirms whether the action took place or not.

```{r data-wrangling, class.source = 'fold-hide'}

# Data importation and cleaning states

ProtestClean <-
  fread("Raw-data/protest_data.csv", na.strings = "", integer64 = "double"
  )[protest == 1
  ][,`:=`(start_date =  paste(startyear,startmonth,startday,sep = "-") |> as_date(),
          end_date =  paste(endyear,endmonth,endday,sep = "-") |> as_date(),
          participants = participants |>  str_to_lower() |> 
            str_remove_all(",") |>
            str_replace_all(" +"," "))
    
    
  # Adding continents based on countries names
  
  ][!country %chin% EuropeMissingCountries, 
    continent := countrycode(country, 
                             origin = "country.name",
                             destination = "continent")
  ][country %chin% EuropeMissingCountries, continent := "Europe"
  
    
  # Extracting values from participants column
  
  ][, c("min","max") := tstrsplit(participants,"-| to ", fixed = FALSE)
  ][participants %like% "between \\d+ and \\d+", 
    `:=`(min = str_match(participants, "between (\\d+) and \\d+")[,2],
         max = str_match(participants, "between \\d+ and (\\d+)")[,2])
  ][, c("min","max") := lapply(.(min, max), 
                               function(x) str_extract(x, "\\d+") |> as.double())
  ][, participants_clean := (min+max)/2
  ][ participants %like% "^\\d+$" & is.na(participants_clean),
     participants_clean :=  as.double(participants)
  ][participants %like% "\\+"& is.na(participants_clean),
    participants_clean := str_remove_all(participants,"\\+|[A-Za-z]") |> as.double()
  ][participants %like% "\\d+s"& is.na(participants_clean),
    participants_clean :=  str_match(participants,"(\\d+)s")[,2] |>  as.double()
  ][participants %like% str_glue("\\d+ ?({KeyWordsAfterNumber})") & 
      is.na(participants_clean),
    participants_clean := str_match(participants,
                                    str_glue("(\\d+) ?({KeyWordsAfterNumber})"))[,2] |>
                          as.double()
  ][participants %like% str_glue("({KeyWordsBeforeNumber}) ?\\d+") &
      is.na(participants_clean),
    participants_clean := str_match(participants,
                                    str_glue("({KeyWordsBeforeNumber}) ?(\\d+)"))[,3] |> 
                          as.double()
  
    
  # Adding after reading the description
  
  ][id == 922006004, participants_clean := 50
  ][id == 6602002005, participants_clean := 2000
    
  
  # Adding the Number of participants by range category
  
  ][, participants_category_clean := 
      fcase( between(participants_clean, 1, 99), "1-99",
             between(participants_clean, 100, 999), "100-999",
             between(participants_clean, 1000, 1999), "1000-1999",
             between(participants_clean, 2000, 4999), "2000-4999",
             between(participants_clean, 5000, 10000), "5000-10000",
             participants_clean > 10000, ">10000",
             default = "Missing")
  ][is.na(participants_category), participants_category :=  participants_category_clean
  ][participants_category == "50-99", participants_category := "1-99"
  ][, participants_range := factor(participants_category,
                                    levels = c("Missing","1-99", "100-999",
                                               "1000-1999",  "2000-4999", 
                                               "5000-10000",">10000"))
    
    
  # Keeping and organizing columns' order
  
  ][, !c("participants","participants_clean","participants_category_clean",
         "participants_category",'region',"startyear","startmonth","startday",
         "endyear","endmonth","endday", "protest","min","max")
  ][, setcolorder(.SD, 
                  union(c('id', 'country', 'ccode', 'year', "continent",
                          'protestnumber','protesterviolence', 'location', 
                          "participants_range","start_date","end_date"),
                        names(.SD)))]



# Changing protester demands and state responses to a longer format

ProtesterStateActionsMelted <-
ProtestClean[, melt(.SD, 
                    id.vars = "id",
                    measure.vars = str_subset(names(.SD),
                                              "^protesterdemand|^stateresponse"),
                    value.name = "actions",
                    variable.factor = FALSE)
  ][!is.na(actions) & actions %like% "\\w"
  ][, variable := fifelse(variable %like% "^protesterdemand", 
                          "protester_demand","state_response")
  ][, `:=` (action_occur = TRUE,
            actions = str_c(variable," - ",actions))
  ][, !c("variable")
  ][, unique(.SD)
  ][, merge(.SD,
            CJ(id = unique(ProtestClean$id),
               actions = unique(actions)),
            by = c("id", "actions"), all = TRUE)
  ][is.na(action_occur), action_occur := FALSE
  ][, c("action_source", "actions") := tstrsplit(actions, split = " - ", fixed = TRUE)
  ][,c("id","action_source","actions","action_occur")]


# Joining protester demands and state responses in a longer format

ProtestCleanTidy <-
ProtestClean[, .SD, .SDcols = !patterns("^protesterdemand|^stateresponse")
  ][, merge(.SD,
            ProtesterStateActionsMelted,
            by = "id", all = TRUE)
  ][, setcolorder(.SD, setdiff(names(.SD), 
                               c("protesteridentity","notes","sources")))
  #Making continent a factor to keep constant the color in plots
  ][, continent := fct_reorder(continent, id, .fun = uniqueN, .desc = TRUE)]


# Show final table

TidyExample <-
ProtestCleanTidy[id == id[1], 
                  c('id', 'country', 'participants_range', 'start_date', 
                    'end_date', 'action_source', 'actions', 'action_occur')
  ][, id := as.character(id)] 

flextable(TidyExample) |>
  border_remove() |>
  bg(bg = "#008000", part = "header") |>
  bold(part = "header") |>
  color(color = "white", part = "header") |>
  border_inner_h(border = officer::fp_border(color = "#dddddd")) |>
  bg(i = (1:nrow(TidyExample))[1:nrow(TidyExample) %% 2 == 0], bg = "#f3f3f3") |>
  autofit()

```

# 2. Data exploration


## 2.1. Which country have more protests?

In next chart we can see that the countries with more protests comes from Europe. Together, they sum 12.6% of protest in the data. Let's focus in those countries and try to answer some questions.

```{r}

plot_chr_count(DATA = ProtestCleanTidy,
               id_var = "id",
               group_var = "country",
               plot_fill_var = "continent",
               plot_title = "Top 20 of Coutries with more Protests",
               plot_max = 20,
               use_cumulative_sum = TRUE)
```

## 2.2. UK Graph Analysis

```{r, fig.dim = c(14,10)}
library(igraph)
library(ggraph)

UKEdges <-
  ProtestCleanTidy[action_occur == TRUE & action_source == "protester_demand"
  ][, setnames(.SD, "actions", "from")
  ][ProtestCleanTidy[action_occur == TRUE & action_source == "state_response",
                     .(to = actions), id],
    on = .(id)
  ][country == "United Kingdom", 
    .(number_protests = uniqueN(id)),
    .(from,to)
  ][,unique(.SD)]

UKNodes <-
ProtestCleanTidy[country == "United Kingdom" & action_occur == TRUE,
                 .(actions,action_source)
  ][, unique(.SD)]


UKgraph <- graph_from_data_frame(d = UKEdges, vertices = UKNodes)

set.seed(1)

ggraph(UKgraph, layout = "with_fr")+
  geom_edge_link(aes(alpha = number_protests,edge_width = number_protests))+
  geom_node_point(aes(color = action_source), size = 8)+
  geom_node_text(aes(label = name), vjust = 2, size = 6)+
  scale_color_brewer(palette = "Dark2")+
  labs(title = "Graph of UK Protests")+
  theme_void()+
  theme(legend.position = "left",
        legend.title = element_text(face = "bold", size = 15),
        legend.text = element_text(size = 15),
        plot.title = element_text(face = "bold", size = 20, hjust = 0.5))

```

# 3. More plots

```{r, fig.dim = c(14,10)}

plot_chr_count(DATA = ProtestCleanTidy,
               id_var = "id",
               group_var = "country",
               plot_fill_var = "continent",
               plot_title = "Total of protests per country",
               plot_max = 40,
               use_cumulative_sum = FALSE)+
  labs(subtitle = " ")+
  facet_wrap(~continent, scales = "free_y")+
  theme(legend.position = "none")

```


As we can see in the chart most of the protests occurs in **`r levels(ProtestCleanTidy$continent)[1]`** and  **`r levels(ProtestCleanTidy$continent)[2]`**.

As this data have information of protests between *`r ProtestCleanTidy[, min(year)]`* and *`r ProtestCleanTidy[, max(year)]`* let's see how the number of protests change each year for each continent.

```{r, fig.align='center', fig.dim = c(10,6)}

continent_count_year_plot <-
ProtestCleanTidy[, .(number_protest = uniqueN(id)),
                  .(continent, year)] |>
  SharedData$new(key = ~continent, group = "Select a continent") |>
  ggplot(aes(year, number_protest, color = continent))+
  geom_line(size = 1)+
  geom_point(size = 2)+
  scale_y_continuous(breaks = breaks_width(50))+
  scale_x_continuous(breaks = breaks_width(2))+
  labs(title = "Number of Protests per Continent and Year",
       x = "Year", y = "Number of Protests", color = "Continent")+
  theme(plot.title = element_text(color = "black", face = "bold"))

  ggplotly(continent_count_year_plot) |>
    highlight(on = 'plotly_click',
              off = 'plotly_doubleclick')

```

At first view, we can see that *2020* isn't a normal year as countries have less protests that year than any other year in the past. That abnormality makes sense when we check that he last date in data is ***`r ProtestCleanTidy[,max(c(start_date,end_date))] |> format("%m/%d/%y")`***.

Checking each individual Continent by clicking over each one we can see:

- **Europe:** Only *`r continent_count_year_plot$data[continent == "Europe",  mean(number_protest > 150)|> percent()]`* of years have more than 150 per year and would be interesting to under why was the people protesting in **1990**, **1996-1997** and **2013-2019**.

- **Asia:** The number of protests per year increased from less than 100 protests per year to 159 between **1997 and 2000**, then between **2005 and 2006** and 3 high picks in **2011**, **2014** and **2019**.

- **Africa:** The number of protest had a pick in **2005**, an increase tendency between **2010 and 2012**, two important picks in **2015** and **2019**.

- **America:** It has to important picks in **2015** and **2019**.

- **Oceania:** It just have two picks of 5 protest in **2001** and **2005**.

```{r, fig.align='center', fig.dim = c(10,6)}

ProtestCleanTidy[continent == "Europe", 
                  .(number_protest = uniqueN(id)),
                  .( country, year)] |>
  ggplot(aes(number_protest))+
  geom_histogram(bins = 20)+
  scale_x_log10(breaks = 2^(1:7) |> round())


EUR_count_year_plot <-
ProtestCleanTidy[continent == "Europe", 
                  .(number_protest = uniqueN(id)),
                  .( country, year)
  ][, number_protest_max := max(number_protest), country
  ][number_protest_max >= 30] |>
  SharedData$new(key = ~country, group = "Select a continent") |>
  ggplot(aes(year, number_protest, color = country))+
  geom_line(size = 1)+
  geom_point(size = 2)+
  scale_y_continuous(breaks = breaks_width(10))+
  scale_x_continuous(breaks = breaks_width(2))+
  labs(title = "Number of Protests per Country and Year",
       x = "Year", y = "Number of Protests", color = "Continent")+
  theme(plot.title = element_text(color = "black", face = "bold"))

  ggplotly(EUR_count_year_plot) |>
    highlight(on = 'plotly_click',
              off = 'plotly_doubleclick')

```